---
title: "The Pretty Friend Penalty"
subtitle: "A Network Model of Beauty, Projection, and Female Social Instability"
author: "Diouana Woman®"
date: "May 20, 2026"
output: 
  pdf_document:
    latex_engine: xelatex
    number_sections: false
  geometry: margin=0.8in
  fontsize: 11pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  fig.width = 7,
  fig.height = 5
)

library(tidyverse)
library(knitr)
library(grid)

round_half_up <- function(x, digits = 0) {
  multiplier <- 10^digits
  floor(x * multiplier + 0.5) / multiplier
}

set.seed(2026)
```

## Abstract

This working paper answers a simple social question: what happens when women hate you because men want you? Using a toy network model, the paper models the destabilization of female friendship networks under uneven distributions of male attention. It argues that beauty functions not merely as an individual attribute but as a  perceived social position. When women are socialized to treat desirability as evidence of worth, another woman's beauty may be misread as access, protection, or rank. The model introduces projection intensity as a function of attention gaps, scarcity beliefs, and self-concept instability. The central claim is simple: sometimes women do not dislike another woman because of who she is; they dislike her because of what they believe her beauty gives her. 

## Introduction
Female friendship is often described through the language of solidarity, sisterhood, and emotional intimacy (Doherty, 2021). Yet this language can obscure the ways beauty, desirability, and male attention quietly reorganize the social field among women. In certain settings, beauty does not merely function as an aesthetic trait. It functions as rank (Mears, 2020). 

This paper asks what happens when one woman's beauty is interpreted by other women as access: access to male attention, protection, softness, status, invitations, or social ease. The question is not whether beauty objectively guarantees these outcomes. Rather, the question is how the belief that beauty confers advantage changes the way women relate to one another. The model developed here treats female friendship as a network. Women are represented as nodes, friendship ties as relational bonds, and projection as a force that weakens those bonds when perceived attention gaps become psychologically meaningful. 

## Related Work
Prior research on "The Femme Fatale Effect" suggests that beauty may become a liability for women when attractiveness, coupled with sexual insecurity, produces distrust rather than social advantage (Sheppard and Johnson, 2019). While this work focuses on workplace judgments of truthfulness, trust, and perceived deservingness of termination, it provides a useful precedent for thinking about beauty as socially ambivalent rather than uniformly beneficial. 

This working paper extends that intuition into informal female social networks. Instead of asking whether attractive women are trusted in professional settings, it asks how beauty destabilizes friendship ties when desirability is perceived as a scarce social resource. The model also draws conceptually from social comparison theory, which suggests that individuals evaluate themselves in relation to others (Caliskan et al., 2024), and self-evaluation maintenance theory, which examines how close relationships become strained when comparison threatens self-concept (Tesser, 1988).

## Conceptual Framework: Erotic Capital Economics
Erotic Capital Economics is a Jungian-derived archetypal framework that examines how women relate to beauty and self-concept through economic behavior. As "Jungian theory posits that reality is psychic reality" (Lambert and Ferns, 2024), rather than treat beauty as a fixed personal attribute, this framework treats beauty as a social and symbolic force that becomes legible through personal context and economic exchange. 

The full Erotic Capital Economics framework contains seven feminine archetypes. For analytical brevity, this working paper uses a subset of four archetypes: the Lover, the Huntress, the Mystic, and the Maiden. These four archetypes were selected because they offer distinct relationships to desirability, social comparison, scarcity, and self-concept stability. In this model, archetypes should not be read as fixed identities or clinical categories. They function as conceptual types that allow the model to simulate how different psychic orientations may respond to unequal distributions of male attention within a female friendship network.

## Archetypal Parameterization
The four archetypes selected for this toy model represent distinct psychic orientations toward beauty, desirability, and social comparison. The Lover treats beauty as currency and sensuality as strategy. The Huntress treats beauty as a tool for self-determined success. The Maiden treats beauty as exploration, softness, and identity formation. The Mystic treats beauty as sacred, intuitive, and spiritually resonant. 

To translate these archetypes into model parameters, each archetype is assigned illustrative values for self-concept stability \(S\), self-perceived desirability security \(D\), beauty salience \(B\), and visibility \(V\). These values are not empirical claims. Rather, they are theoretical assumptions used to demonstrate the logic of the model. Self-concept stability \(S\) measures how internally anchored an archetype is. Self-perceived desirability security \(D\) measures whether desirability is believed to be abundant rather than scarce. Beauty salience \(B\) represents how strongly beauty is socially legible in a given setting, while visibility \(V\) represents how likely the archetype is to be noticed within the room. This paper uses an equal weighting of beauty salience and visibility in the male attention function. 

This reflects a core premise of Erotic Capital Economics: desirability is not reducible to physical features. Aura, social legibility, posture, styling, and the ability to become visible within a room may shape attention as powerfully as conventional beauty (Hakim, 2010).

```{r archetype-parameters}
archetypes <- tibble::tribble(
  ~archetype,  ~S,   ~D,   ~B,   ~V,
  "Lover",     0.85, 0.90, 0.90, 0.85,
  "Huntress",  0.80, 0.75, 0.75, 0.75,
  "Maiden",    0.50, 0.40, 0.65, 0.60,
  "Mystic",    0.80, 0.85, 0.70, 0.55
) %>%
  mutate(
    M = 0.50 * B + 0.50 * V
  )

archetype_colors <- c(
  "Lover" = "#FF5B8A",
  "Huntress" = "#49C5D6",
  "Maiden" = "#F2E86D",
  "Mystic" = "#B07AE8"
)

knitr::kable(
  archetypes %>%
    mutate(
      across(c(S, D, B, V, M), ~ round_half_up(.x, 2))
    ),
  digits = 2,
  caption = "Illustrative archetypal parameters for the Pretty Friend Penalty model."
)
```

## Formal Model

Let each woman in a friendship group be represented as a node \(i\) in a directed network \(G\). A directed tie \(F_{j \to i}\) represents woman \(j\)'s relational openness toward woman \(i\). This direction matters because projection is often asymmetric: one woman may experience another woman as a threat even when the feeling is not reciprocated.

Male attention is modeled as a weighted function of beauty salience and visibility:

\[
M_i = \beta_B B_i + \beta_V V_i
\]

where \(B_i\) is beauty salience, \(V_i\) is visibility, and \(\beta_B + \beta_V = 1\). In this model, \(\beta_B = 0.50\) and \(\beta_V = 0.50\).

The attention gap from woman \(j\)'s perspective toward woman \(i\) is:

\[
\Delta M_{j \to i} = \max(M_i - M_j, 0)
\]

This expression captures only positive attention gaps. If woman \(i\) receives more male attention than woman \(j\), the gap is positive. If woman \(i\) receives equal or less attention, the gap is set to zero.

Projection from woman \(j\) onto woman \(i\) is modeled as:

\[
P_{j \to i} = \Delta M_{j \to i}(1-D_j)(1-S_j)
\]

where \(D_j\) is woman \(j\)'s desirability security and \(S_j\) is woman \(j\)'s self-concept stability. The terms \(1-D_j\) and \(1-S_j\) represent complementary psychological states: desirability insecurity and self-concept instability. Projection therefore increases when the observer experiences desirability as scarce and has a less stable self-concept.

Friendship stability then evolves according to:

\[
F_{j \to i}(t+1) = \max\{0,\; F_{j \to i}(t) - \alpha P_{j \to i}\}
\]

where \(t\) represents the interaction period and \(\alpha\) is the sensitivity parameter. When \(\alpha\) is low, the friendship network is resilient and projection produces only minor relational damage. When \(\alpha\) is high, the network is fragile and even moderate projection may sharply reduce friendship stability.

## Simulation
To demonstrate the model, this section simulates directed projection among the four selected archetypes. The initial friendship tie is set to \(F_0 = 1\), representing full relational openness. The sensitivity parameter is set to \(\alpha = 2.5\), allowing projection to produce visible but not automatically catastrophic friendship decline.

```{r projection-simulation, results='asis'}
alpha <- 2.5
F0 <- 1

edges <- tidyr::expand_grid(
  from = archetypes$archetype,
  to = archetypes$archetype
) %>%
  filter(from != to) %>%
  left_join(
    archetypes %>% select(to = archetype, M_to = M),
    by = "to"
  ) %>%
  left_join(
    archetypes %>% select(
      from = archetype,
      M_from = M,
      D_from = D,
      S_from = S
    ),
    by = "from"
  ) %>%
  mutate(
    attention_gap = pmax(M_to - M_from, 0),
    desirability_insecurity = 1 - D_from,
    self_concept_instability = 1 - S_from,
    projection = attention_gap * desirability_insecurity * self_concept_instability,
    friendship_next = pmax(0, F0 - alpha * projection)
  ) %>%
  arrange(desc(projection))

cat("\\begingroup\n")
cat("\\small\n")
cat("\\setlength{\\tabcolsep}{5pt}\n")
cat("\\renewcommand{\\arraystretch}{1.08}\n")

knitr::kable(
  edges %>%
    mutate(
      attention_gap = round_half_up(attention_gap, 3),
      desirability_insecurity = round_half_up(desirability_insecurity, 2),
      self_concept_instability = round_half_up(self_concept_instability, 2),
      projection = round_half_up(projection, 3),
      friendship_next = round_half_up(friendship_next, 3)
    ) %>%
    select(
      observer = from,
      target = to,
      attention_gap,
      desirability_insecurity,
      self_concept_instability,
      projection,
      friendship_next
    ),
  col.names = c(
    "Observer",
    "Target",
    "$\\Delta M$",
    "$1-D$",
    "$1-S$",
    "$P$",
    "$F(t+1)$"
  ),
  digits = 3,
  escape = FALSE,
  caption = "Directed projection and resulting friendship stability between archetypes."
)

cat("\\endgroup\n")
```


```{r network-diagram}
edge_threshold <- 0.005

node_pos <- archetypes %>%
  select(name = archetype, M) %>%
  mutate(
    x = case_when(
      name == "Lover" ~ 0,
      name == "Huntress" ~ 1.45,
      name == "Maiden" ~ 0,
      name == "Mystic" ~ -1.45
    ),
    y = case_when(
      name == "Lover" ~ 1.35,
      name == "Huntress" ~ 0,
      name == "Maiden" ~ -1.35,
      name == "Mystic" ~ 0
    ),
    m_x = case_when(
      name == "Lover" ~ x,
      name == "Huntress" ~ x + 0.55,
      name == "Maiden" ~ x,
      name == "Mystic" ~ x - 0.55
    ),
    m_y = case_when(
      name == "Lover" ~ y + 0.38,
      name == "Huntress" ~ y - 0.02,
      name == "Maiden" ~ y - 0.42,
      name == "Mystic" ~ y - 0.02
    )
  )

plot_edges <- edges %>%
  filter(projection >= edge_threshold) %>%
  left_join(
    node_pos %>% select(from = name, x_from = x, y_from = y),
    by = "from"
  ) %>%
  left_join(
    node_pos %>% select(to = name, x_to = x, y_to = y),
    by = "to"
  ) %>%
  mutate(
    dx = x_to - x_from,
    dy = y_to - y_from,
    dist = sqrt(dx^2 + dy^2),
    ux = dx / dist,
    uy = dy / dist,
    node_radius = 0.28,
    x_start = x_from + ux * node_radius,
    y_start = y_from + uy * node_radius,
    x_end = x_to - ux * node_radius,
    y_end = y_to - uy * node_radius,
    edge_label = paste0("P = ", sprintf("%.3f", round_half_up(projection, 3))),
    label_x = case_when(
      from == "Maiden" & to == "Lover" ~ -0.38,
      from == "Maiden" & to == "Huntress" ~ 0.86,
      from == "Mystic" & to == "Lover" ~ -0.98,
      from == "Huntress" & to == "Lover" ~ 0.98,
      from == "Mystic" & to == "Huntress" ~ 0,
      TRUE ~ (x_start + x_end) / 2
    ),
    label_y = case_when(
      from == "Maiden" & to == "Lover" ~ 0.00,
      from == "Maiden" & to == "Huntress" ~ -0.82,
      from == "Mystic" & to == "Lover" ~ 0.90,
      from == "Huntress" & to == "Lover" ~ 0.90,
      from == "Mystic" & to == "Huntress" ~ 0.22,
      TRUE ~ (y_start + y_end) / 2
    )
  )

ggplot() +
  geom_segment(
    data = plot_edges,
    aes(
      x = x_start,
      y = y_start,
      xend = x_end,
      yend = y_end,
      linewidth = projection
    ),
    arrow = arrow(length = unit(3.5, "mm"), type = "closed"),
    color = "black",
    alpha = 0.78,
    lineend = "round"
  ) +
  geom_point(
    data = node_pos,
    aes(x = x, y = y, fill = name),
    shape = 21,
    size = 18,
    color = "white",
    stroke = 1.2,
    alpha = 0.95
  ) +
  geom_text(
    data = node_pos,
    aes(x = x, y = y, label = name),
    color = "black",
    fontface = "bold",
    size = 3.5
  ) +
  geom_text(
    data = node_pos,
    aes(
      x = m_x,
      y = m_y,
      label = paste0("M = ", sprintf("%.2f", round_half_up(M, 2)))
    ),
    color = "gray20",
    size = 3.0
  ) +
  geom_text(
    data = plot_edges,
    aes(x = label_x, y = label_y, label = edge_label),
    color = "black",
    size = 3.0
  ) +
  scale_fill_manual(values = archetype_colors) +
  scale_linewidth_continuous(range = c(0.45, 1.35)) +
  coord_equal(xlim = c(-2.25, 2.25), ylim = c(-1.9, 1.9)) +
  labs(
    title = "Directed Projection Field",
    subtitle = "Arrows move from observer to target; projection scores appear beside each tie"
  ) +
  theme_void() +
  theme(
    legend.position = "none",
    plot.title = element_text(hjust = 0.5, face = "bold", size = 15),
    plot.subtitle = element_text(hjust = 0.5, size = 9.5, color = "gray30")
  )
```

The directed network visualization shows that projection does not move evenly across the friendship group. Projection is strongest when an archetype with lower desirability security and lower self-concept stability observes another archetype receiving greater male attention. This is why direction matters: the Lover may not project onto the Maiden, even if the Maiden projects onto the Lover. The model therefore captures a common asymmetry in female social life, where one woman experiences rivalry while the other experiences mere coexistence (hooks, 1986).

## Discussion
The simulation suggests that beauty does not destabilize female friendship by itself. Rather, instability emerges when beauty is interpreted as social rank. Uneven male attention becomes socially meaningful when women are taught to experience desirability as scarce and to interpret another woman’s visibility as their own diminishment (Merrie et al., 2025).

The Pretty Friend Penalty therefore describes a relational penalty imposed on women whose beauty becomes legible as access. The beautiful woman is not merely seen as beautiful; she is imagined to possess protection, softness, status, invitations, and ease. Whether or not she actually possesses these advantages is secondary to the belief that she does.

This model also clarifies why accusations such as “she thinks she is special” often appear around beautiful or highly visible women. The accusation displaces the observer’s projection onto the target. Instead of naming envy, comparison, or desirability insecurity, the target is recoded as arrogant, attention-seeking, or socially threatening (Sheppard and Johnson, 2019).

Within the Erotic Capital Economics framework, the archetypes respond differently to this dynamic. The Lover may receive attention without interpreting desirability as scarce. The Huntress may recognize the competitive structure but remain strategically self-directed. The Mystic may be less reactive because her sense of value is internally and spiritually sourced. The Maiden, still emerging into her self-concept, may be more vulnerable to comparison when desirability appears unevenly distributed.

## Conclusion
This working paper introduced the Pretty Friend Penalty as a toy network model of beauty, projection, and female social instability. The model argues that female friendship networks destabilize not because women are inherently jealous, but because desirability is socially overvalued and intellectually trivialized. When male attention is treated as proof of worth, another woman’s beauty may be experienced as symbolic deprivation.

The model is intentionally simple. Its purpose is not to make empirical claims about all women, but to formalize a recurring social pattern: sometimes women do not dislike another woman because of who she is; they dislike her because of what they believe her beauty gives her.

Future versions of this model may incorporate the full seven-archetype Erotic Capital Economics framework, empirical survey data, and longitudinal friendship dynamics. For now, the Pretty Friend Penalty offers a first mathematical language for understanding how beauty, projection, and social hierarchy move through female friendship networks.

## References
Caliskan, F., Idug, Y., Uvet, H., Gligor, N., & Kayaalp, A. (2024). Social comparison theory: A review and future directions. Psychology & Marketing, 41(11), 2823–2840. https://doi.org/10.1002/mar.22087

Doherty, E. F. (2021). “It’s just like a break-up”: A qualitative analysis of turning points in female friendships in early to middle adulthood. Communication Quarterly, 69(1), 43–66. https://doi.org/10.1080/01463373.2021.1877163

Hakim, C. (2010). Erotic Capital. European Sociological Review, 26(5), 499–518. https://doi.org/10.1093/esr/jcq014

hooks, b. (1986). Sisterhood: Political Solidarity between Women. Feminist Review, (23), 125–138. https://doi.org/10.2307/1394725

Lambert, A., & Ferns, G. (2024). Into the depths of the feminine: A Jungian perspective on postfeminist working life. Human Relations, 77(10), 1534–1562. https://doi.org/10.1177/00187267231199650

Mears, A. (2020). Very Important People: Status and Beauty in the Global Party Circuit. Princeton University Press. https://doi.org/10.1515/9780691189895

Merrie, L. A., Krems, J. A., & Byrd-Craven, J. (2025). Who is a mating rival? Women track other women’s intent and capacity to compete for mates. Evolutionary Behavioral Sciences. https://doi.org/10.1037/ebs0000372

Sheppard, L. D., & Johnson, S. K. (2019). The Femme Fatale Effect: Attractiveness is a Liability for Businesswomen’s Perceived Truthfulness, Trust, and Deservingness of Termination. Sex Roles, 81(11–12), 779–796. https://doi.org/10.1007/s11199-019-01031-1

Tesser, A. (1988). Toward a Self-Evaluation Maintenance Model of Social Behavior. In Advances in Experimental Social Psychology (Vol. 21, pp. 181–227). Elsevier. https://doi.org/10.1016/S0065-2601(08)60227-0